Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries

### 2024-04-21 paper.qmd
chi squared test works

```{python}
# Summarize missing data for InjuryRecord.csv
injury_missing_summary = injury.isnull().sum()
injury_missing_summary = injury_missing_summary[injury_missing_summary > 0]

# Summarize missing data for PlayList.csv
playlist_missing_summary = plays.isnull().sum()
playlist_missing_summary = playlist_missing_summary[playlist_missing_summary > 0]

# Convert the Series to a DataFrame and transpose it for horizontal layout
injury_missing_df = injury_missing_summary.to_frame().transpose()
playlist_missing_df = playlist_missing_summary.to_frame().transpose()

# Calculate the number of missing data in each column divided by the number of rows for InjuryRecord
missing_data_ratio_injury = (28) / len(injury)

# Calculate the number of missing data in each column divided by the number of rows for PlayList
missing_data_ratio_playlist = (16910 + 18691 + 367)/ len(plays)

# Display the DataFrames
print('InjuryRecord.csv')
print(injury_missing_df) 
print('Ratio:', missing_data_ratio_injury)
print('PlayList.csv')
print(playlist_missing_df)
print('Ratio:', missing_data_ratio_playlist)
```

From `InjuryRecord.csv`, there are only 28 values missing in the `PlayKey` column from the original 105. At 27%, there is a significant amount of missing data entries. However, it would be difficult to fill in the missing entries without using an external dataset. This variable is also within the `PlayList.csv` file so it can be used to fill in the data in `InjuryRecord.csv` (if applicable).

In `PlayList.csv`, 13% of the data is missing, mainly in `StadiumType`, `Weather`, and `PlayType`. Similar to before, it is difficult to fill in data without an external resource. For now, the models and analysis will use the data available in the files.

@@

# Save the modified dataframe if needed
injury.to_csv('data/InjuryRecord.csv', index=False)
```
# Modeling

@@

# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(8, 4))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()

@@

# Create a contingency table of injuries by surface type
contingency_table = pd.crosstab(plays['InjuryOccurred'], injury['Surface'])

@@

print(f"Chi-Square Statistic: {chi2}, p-value: {p}")
```

>>>>

```{python}
# Summarize missing data for InjuryRecord.csv
injury_missing_summary = injury.isnull().sum()
injury_missing_summary = injury_missing_summary[injury_missing_summary > 0].rename('InjuryRecord')

# Summarize missing data for PlayList.csv
playlist_missing_summary = plays.isnull().sum()
playlist_missing_summary = playlist_missing_summary[playlist_missing_summary > 0].rename('PlayList')

# Calculate the ratio of missing data to total entries for each dataset
missing_data_ratio_injury = (injury_missing_summary.sum() / len(injury)) * 100
missing_data_ratio_playlist = (playlist_missing_summary.sum() / len(plays)) * 100

# Create a DataFrame from the missing data summaries
missing_data_summary = pd.DataFrame({
    'InjuryRecord.csv': injury_missing_summary,
    'PlayList.csv': playlist_missing_summary
})

# Fill NaN values with zero and convert all numeric counts to integers
missing_data_summary = missing_data_summary.fillna(0).astype(int)

# Convert all numeric counts to integers (ensures whole numbers)
missing_data_summary = missing_data_summary.astype(int)

# Add the ratio of missing data as a new row
missing_data_summary.loc['Missing Data Ratio (%)'] = [f"{missing_data_ratio_injury:.2f}%", f"{missing_data_ratio_playlist:.2f}%"]

# Transpose the DataFrame for better readability
missing_data_summary = missing_data_summary.T

# Rename index to indicate it's the file name and adjust column names if needed
missing_data_summary.index.name = 'File Name'

# Display the DataFrame neatly
print(missing_data_summary)
```

From `InjuryRecord.csv`, there are only 28 values missing in the `PlayKey` column from the original 105. At 27%, there is a significant amount of missing data entries. However, it would be difficult to fill in the missing entries without using an external dataset. This variable is also within the `PlayList.csv` file so it can be used to fill in the data in `InjuryRecord.csv` (if applicable). In `PlayList.csv`, 13% of the data is missing, mainly in `StadiumType`, `Weather`, and `PlayType`. Similar to before, it is difficult to fill in data without an external resource. For now, the models and analysis will use the data available in the files.

@@

# Save the modified dataframe if needed
injury.to_csv('data/mod_InjuryRecord.csv', index=False)
```

For the missing entries in `InjuryRecord.csv`, we can fill in the missing data by formatting the missing values based on the `GameID`. This step allows us to locate the data between each file and determine important factors that contribute to injuries.

```{python}
# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(8, 4))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()
```

This graph models all of the different stadium types that are in the file. Notice how there are many repeats that can be combined. In order to reduce repetitiveness, these values can be cleaned by combining like-values.

```{python}
# A quick view at the 'StadiumType' unique values to understand the variations
stadium_types = plays['StadiumType'].unique()

stadium_type_mapping = {
    'Retr. Roof-Closed': 'Retr. Roof - Closed',
    'Retr. Roof-Open': 'Retr. Roof - Open',
    'Outdoor': 'Outdoors',
    'Ourdoor': 'Outdoors',
    'Oudoor': 'Outdoors',
    'Outddors': 'Outdoors',
    'Outdor': 'Outdoors',
    'Outside': 'Outdoors',
    'Heinz Field': 'Outdoors',  # Assuming Heinz Field is always outdoor
    'Cloudy': 'Outdoors',  # Assuming 'Cloudy' is an outdoor condition and not a stadium type
    'Indoor': 'Indoors',
    'Domed': 'Dome',
    'Domed, closed': 'Dome, closed',
    'Domed, Open': 'Dome, closed',
    'Closed Dome': 'Dome, closed',
    'Retr. Roof Closed': 'Retr. Roof - Closed'
    # More mappings can be added here based on the actual data
}

# Standardize the 'StadiumType' column
plays['StadiumType'] = plays['StadiumType'].replace(stadium_type_mapping)

# Calculate the new frequency of each stadium type
cleaned_stadium_freq = plays['StadiumType'].value_counts()

# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(8, 4))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()
```

# Modeling

@@

@@

# Ensure 'Surface' column is in the PlayList dataset and check data
if 'FieldType' not in plays.columns:
    print("Error: 'FieldType' column is missing from PlayList dataset.")
else:
    # Create a contingency table of injuries by surface type
    contingency_table = pd.crosstab(plays['InjuryOccurred'], plays['FieldType'])

@@

    print(f"Chi-Square Statistic: {chi2:.3f}, p-value: {p:.3f}")
```
